### 2016-10-19 latex/Adaptive_distance_metric_notes.tex
sync with emails

\documentclass[]{article}

\renewcommand{\baselinestretch}{1.2}
\renewcommand{\v}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\mc}{\mathcal}

\def\st{{\rm s.t.}}

\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage[usenames,dvipsnames]{color}
\usepackage{fullpage,graphicx,amssymb,amsmath}
\usepackage{enumerate}
\usepackage{subfigure}

\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{assumption}{Assumption}
\newtheorem{conjecture}{Conjecture}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newcommand{\hs}[1]{\hspace{#1}}
\newcommand{\vs}[1]{\vspace{#1}}
\newcommand{\mnorm}[1]{||{#1}||_{\infty}}
\newcommand{\pf}{\textbf{Proof} \indent}
\newcommand{\qed}{\hfill $\Box$}
\newcommand{\argmin}{\rm argmin}
\newcommand{\proof}{\pf}

\newcommand{\Cvar}{\mathbb{CVAR}} % Conditional value-at-risk
%\newcommand{\Cvar}{\mathbb{CV}@\mathbb{R}} % Conditional value-at-risk
\newcommand{\Var}{\mathbb{VAR}} % Value-at-risk
%\newcommand{\Var}{\mathbb{V}\text{a}\mathbb{R}} % Value-at-risk
%\newcommand{\Var}{\mathbb{V}@\mathbb{R}} % Value-at-risk

\newcommand{\cone}{{\rm cone}} % Cone
\newcommand{\conv}{{\rm conv}} % Convex hull
\newcommand{\clconv}{{\rm clconv}} % Closed convex hull
\newcommand{\Cov}{{\rm Cov}} % Covariance
\newcommand{\D}{\mathbb{D}} % Distance metric
\newcommand{\E}{\mathbb{E}} % Expectation
\renewcommand{\P}{\mathbb{P}} % Probability
\renewcommand{\Re}{\mathbb{R}} % Real numbers
\renewcommand{\S}{\mathcal{S}} % Set of scenarios
\newcommand{\Z}{\mathbb{Z}} % Integers

\newcommand{\vell}{\boldsymbol{\ell}}
\newcommand{\valpha}{\boldsymbol{\alpha}}
\newcommand{\vbeta}{\boldsymbol{\beta}}
\newcommand{\vgamma}{\boldsymbol{\gamma}}
\newcommand{\vGamma}{\boldsymbol{\Gamma}}
\newcommand{\vdelta}{\boldsymbol{\delta}}
\newcommand{\vlambda}{\boldsymbol{\lambda}}
\newcommand{\vLambda}{\boldsymbol{\Lambda}}
\newcommand{\vxi}{\boldsymbol{\xi}}
\newcommand{\vpi}{\boldsymbol{\pi}}

\newcommand{\xor}{\underline{\vee}}
\newcommand{\Xor}{\underline{\bigvee}}

\title{Adaptive Distance Metric Selection for Supervised Classification}
\author{Krishnan Kumaran and Dimitri J. Papageorgiou \\
{\small Corporate Strategic Research}\\
{\small ExxonMobil Research and Engineering Company}\\
{\small 1545 Route 22 East, Annandale, NJ 08801 USA}\\
{\small dimitri.j.papageorgiou@exxonmobil.com} \\
}

\begin{document}

\maketitle

\section{Introduction and motivation}

Clustering (also known as unsupervised classification) and Classification (commonly understood to be supervised, i.e., based on training data) are common ways of performing data analysis to perform a range of functions like anomaly detection and diagnosis, data segmentation and model development/refinement. Consequently, there is a large body of research on these topics (see for example [1] and references therein for a survey of currently used methods) offering different solutions ranging from K-means, spectral methods and other more complex methods for Clustering, as well as Support Vector Machines, Artificial Neural Networks (ANN), their non-linear Kernel-based variants and others for Classification.

All clustering methods require a similarity/distance metric between data points. While there are typically a few reasonable choices for most data, the results can depend strongly on this choice. The main goal of this project is to extend existing clustering work (e.g., [2,3,4]) to a semi-supervised classification method that is capable of learning the optimal distance/similarity metric directly from training data. The clustering problem can be formulated as an optimization problem (a mixed-integer linear program in its simplest form) that can be hard to solve for realistic data sizes. This approach using an adaptive distance metric has the potential to provide results of better quality than state-of-the-art techniques like Support Vector Machines.

\section{Problem statement}

Suppose we are given $N$ points $\v{v}_i \in \Re^p$ with class membership $\mc{C}_i$ for $i \in \mc{N} = \{1,\dots,N\}$.
Our goal is to find a distance metric $\D: \Re^p \mapsto \Re$ such that the following conditions hold:
\begin{enumerate}
\item $\D(\v{v}_i,\v{v}_j) = \v{a}^{\top}(\v{v}_i - \v{v}_j) + (\v{v}_i - \v{v}_j)^{\top} \v{B}(\v{v}_i - \v{v}_j) \quad \forall i,j \in \mc{N}$, for some $\v{a} \in \Re^p$ and $\v{B} \in \Re^{p \times p}$ and symmetric
\item $\D(\v{v}_i,\v{v}_j) \geq 0 \quad \forall i,j \in \mc{N}$
\item $\min_{j \in \mc{C}_i} \D(\v{v}_i,\v{v}_j) < \min_{k \notin \mc{C}_i} \D(\v{v}_i,\v{v}_j) \quad \forall i \in \mc{N}$
\end{enumerate}
Note that this is not a true ``distance metric'' since symmetry and the triangle property are not enforced.
The first condition could be extended to include higher order terms.
Let $\vdelta_{ij} = \v{v}_i - \v{v}_j$ for all $i,j \in \mc{N}$.
Finding such a distance metric can be expressed as the following optimization problem:

\begin{subequations} \label{model:1}
\begin{alignat}{4}
\max_{\lambda,\v{a},\v{B},\v{d}}~~& \lambda &&  \\
\st~~& \min_{j \in \mc{C}_i} d_{ij} + \lambda \leq \min_{k \notin \mc{C}_i} d_{ik} & & \qquad \forall i \in \mc{N} \label{eq:separation_requirement} \\
    & d_{ij} = \v{a}^{\top}\vdelta_{ij} + \vdelta_{ij}^{\top}\v{B}\vdelta_{ij} && \qquad \forall i,j \in \mc{N} \\
    & d_{ij} \geq 0 && \qquad  \forall i,j \in \mc{N} \\
    & \v{a} \in \Re^p && \\ 
    & \v{B} \in \Re^{p \times p}, \text{symmetric} &&
\end{alignat}
\end{subequations}

Let $\lambda_i = \left[ \min_{k \notin \mc{C}_i} d_{ik} \right] - \left[ \min_{j \in \mc{C}_i} d_{ij} \right]$ be the difference between the separation/distance from point $i$ to its nearest nonneighbor and the separation/distance from point $i$ to its nearest neighbor.  Ideally, we would like $\lambda_i > 0$ for all $i$.  However, this might not be possible when restricted to a set of distance functions.
In formulation \eqref{model:1}, we would like to maximize the minimum separation over all points as can be seen by re-writing constraint \eqref{eq:separation_requirement} as
$$
\lambda \leq \lambda_i \qquad \forall i \in \mc{N}
$$ 
and noting that the objective function is to maximize $\lambda$.

\subsection{Potential loss functions}

Let $L(\vlambda) = L(\lambda_1,\dots,\lambda_n)$ be a loss function based on the separation of each point $i$. 
Let $f = -L$ be the negative loss function.
In formulation \eqref{model:1}, we are interested in maximizing the function 
$$
f(\vlambda) = -L(\vlambda) = \min_{i \in \mc{N}} \lambda_i~.
$$
Of course, other loss functions are possible and should potentially be considered.

Option 1: Assume $f$ is linear and separable: 
\begin{equation}
f(\vlambda) = \sum_{i \in \mc{N}} f_i(\lambda_i)
\end{equation}

Option 2: Assume $f$ is piecewise linear increasing and separable: 
\begin{equation}
f^{\textrm{PWL-Cont}}(\lambda_i) = \left\{
\begin{array}{ll}
\alpha \lambda_i & \lambda_i \geq 0 \\
\beta \lambda_i & o.w. \\
\end{array}
\right.
\textrm{with } \alpha \leq \beta
\end{equation}

Option 3: Assume $f$ is piecewise linear, increasing up to a  and separable: 
\begin{equation}
f^{\textrm{PWL-DisCont}}(\lambda_i) = \left\{
\begin{array}{ll}
\alpha \lambda_i & \lambda_i \geq 0 \\
\beta & o.w. \\
\end{array}
\right.
\textrm{with }
\end{equation}

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.7]{Loss_functions_concatenated.png}
%\vspace{-0.25in}
\caption{Potential loss functions}
\label{fig:loss_functions_concatenated}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
%\includegraphics[width=5.0in,height=2.5in]{Circles_before_and_after_basic_step.png}
\includegraphics[scale=0.7]{sigmoid_from_wikipedia.png}
%\vspace{-0.25in}
\caption{Sigmoid function. [Source: wikipedia]}
\label{fig:sigmoid}
\end{center}
\end{figure}

\subsection{Imposing a $K>1$ nearest neighbor condition}

Let $d_{i(j)}$ denote the distance of the $j$th closest point to point $i \in \mc{N}$, i.e.,
$d_{i(1)} \leq d_{i(2)} \dots\leq d_{i(|\mc{C}_i|)}$. 
Then, inequality \eqref{eq:separation_requirement} can be rewritten as 
\begin{equation}
d_{i(1)} + \lambda \leq d_{ik} \forall i \in \mc{N}, k \notin \mc{C}_i
\end{equation}

A more restrictive requirement 
\begin{equation}
d_{i(t)} + \lambda \leq d_{ik} \forall i \in \mc{N}, k \notin \mc{C}_i
\end{equation}

The above model can be generalized so that the requirement in inequality \eqref{eq:separation_requirement} holds for the $K$ closest points $j \in \mc{C}_i$, and

\newpage
%\section{A linear programming approach to a Feasibility Variant of Model \eqref{model:1}}
\section{A linear programming approach to determine the existence of a feasible distance metric}

Let $\mc{F}$ denote the feasible region associated with the distance metric parameters, i.e.,
\begin{equation}
\mc{F} = \{ (\v{a},\v{B},\v{d}) \in \Re^p \times \Re^{p \times p} \times \Re_+^{N \times N} : \v{B}~\text{symmetric}, d_{ij} = \v{a}^{\top}\vdelta_{ij} + \vdelta_{ij}^{\top} \v{B}\vdelta_{ij} \quad \forall i,j \in \mc{N} \}
\end{equation}
or
\begin{subequations} \label{set:feasible_region_distance_metric}
\begin{alignat}{4}
\mc{F} = \Bigg\{ \v{a},\v{B},\v{d} : &  &&  \\
    & d_{ij} = \sum_{k=1}^p \delta_{ijk} a_{k} + \sum_{k=1}^p \delta_{ijk}^2 b_{kk} + 2 \sum_{k=1}^{p-1} \sum_{\ell=k+1}^p  \delta_{ijk} \delta_{ij\ell} b_{k\ell} && \qquad \forall i,j \in \mc{N} \label{eq:distance_metric_def} \\
    & 0 \leq d_{ij} \leq 1 && \qquad \forall i,j \in \mc{N} \\
    & \v{a} \in \Re^p && \\ 
    & b_{k\ell} \in \Re && \qquad \forall k=1,\dots,p, \ell=k,\dots,p \Bigg\}
\end{alignat}
\end{subequations}

The existence of a distance metric satisfying conditions 1-3 can be expressed as the following feasibility (nonlinear optimization) problem, where $\epsilon > 0$ is a given parameter:
%\begin{subequations} \label{model:1_feasibility}
%\begin{alignat}{4}
%\max_{\v{a},\v{B},\v{d}}~~& 0 &&  \\
%\st~~& \min_{j \in \mc{C}_i} d_{ij} + \epsilon \leq \min_{k \notin \mc{C}_i} d_{ik} & & \qquad \forall i \in \mc{N} \label{eq:separation_requirement_eps} \\
%    & d_{ij} = \v{a}^{\top}\vdelta_{ij} + \vdelta_{ij}^{\top} \v{B}\vdelta_{ij} && \qquad \forall i,j \in \mc{N} \\
%    & d_{ij} \geq 0 && \qquad  \forall i,j \in \mc{N} \\
%    & \v{a} \in \Re^p && \\ 
%    & \v{B} \in \Re^{p \times p}, \text{symmetric} &&
%\end{alignat}
%\end{subequations}
\begin{subequations} \label{model:1_feasibility}
\begin{alignat}{4}
\max_{\v{a},\v{B},\v{d}}~~& 0 &&  \\
\st~~& \min_{j \in \mc{C}_i} d_{ij} + \epsilon \leq \min_{k \notin \mc{C}_i} d_{ik} & & \qquad \forall i \in \mc{N} \label{eq:separation_requirement_eps} \\
    & (\v{a},\v{B},\v{d}) \in \mc{F} &&
\end{alignat}
\end{subequations}
Note that if we constrain the distances to be bounded (e.g., $d_{ij} \in [0,1]$), problem \eqref{model:1_feasibility} is not guaranteed to be feasible for an arbitrary choice of $\epsilon$. 
To see this, let $\epsilon > 0$ and consider an example with three points and two classes.  Point 1 is located at (0,0), point 2 at (1,1), and point 3 at $(\epsilon/2,\epsilon/2)$.  Points 1 and 2 belong to class 1, while point 3 belongs to class 2. There is no way to create a distance metric that guarantees that $d_{13} > \epsilon$ and $d_{23} > \epsilon$.

There are several ways to proceed.

\subsection{Multi-Objective approach}
Let $W$ be a positive scalar.  
We could solve the following LP that maximizes both $\epsilon$ and the minimum distance variables $t_i$:
\begin{subequations} \label{model:multi_objective_lp_1}
\begin{alignat}{4}
\max_{\v{a},\v{B},\v{d},\epsilon,\v{t}}~~& W\epsilon + \sum_{i \in \mc{N}} t_i &&  \\
\st~~& t_i + \epsilon \leq d_{ik} && \qquad \forall i \in \mc{N}, k \in \mc{N}\backslash\mc{C}_i \label{eq:mo_lp_separation_constr_t} \\
    & t_i \leq d_{ij} && \qquad \forall i \in \mc{N}, j \in \mc{C}_i \label{eq:mo_lp_dual_constr_t} \\
    & t_i \geq 0 & & \qquad \forall i \in \mc{N} \\
    & \eqref{set:feasible_region_distance_metric} & & 
\end{alignat}
\end{subequations}
\begin{proposition}
Problem \eqref{model:multi_objective_lp_1} is always feasible.
\end{proposition}
\proof The origin is always a feasible solution.  That is, set $\v{a} = \v{0},\v{B} = \v{0},\v{d} = \v{0},\epsilon = 0,\v{t} = \v{0}$.
\qed

\subsection{Iterative LP approach}

First solve the following problem to determine the maximum non-neighbor separation $\epsilon$ that is possible. 
\begin{subequations} \label{model:epsilon_lp_1}
\begin{alignat}{4}
\max_{\v{a},\v{B},\v{d},\epsilon}~~& \epsilon &&  \\
\st~~& \epsilon \leq d_{ik} && \qquad \forall i \in \mc{N}, k \in \mc{N}\backslash\mc{C}_i \label{eq:epsilon_lp_separation_constr_t} \\
    & \eqref{set:feasible_region_distance_metric} & & 
\end{alignat}
\end{subequations}
Let $\epsilon^{\max}$ be the maximum non-neighbor separation, i.e., the optimal value of $\epsilon$ in \eqref{model:epsilon_lp_1}.
 Note that this problem is identical to problem \eqref{model:multi_objective_lp_1} after omitting the $\v{t}$ decision variables.  Hence, it is guaranteed to be feasible.

Given $\epsilon \in (0,\epsilon^{\max}]$, model \eqref{model:1} can be expressed as a linear programming feasibility problem:
\begin{subequations} \label{model:big_lp_1}
\begin{alignat}{4}
\max_{\v{a},\v{B},\v{d},\v{t}}~~& \sum_{i \in \mc{N}} t_i &&  \\
\st~~& t_i + \epsilon \leq d_{ik} && \qquad \forall i \in \mc{N}, k \in \mc{N}\backslash\mc{C}_i \label{eq:big_lp_separation_constr_t} \\
    & t_i \leq d_{ij} && \qquad \forall i \in \mc{N}, j \in \mc{C}_i \label{eq:big_lp_dual_constr_t} \\
    & t_i \geq 0 & & \qquad \forall i \in \mc{N} \\
    & \eqref{set:feasible_region_distance_metric} & & 
\end{alignat}
\end{subequations}
For $\epsilon \in (0,\epsilon^{\max}]$, model \eqref{model:big_lp_1} is always feasible since the origin $\v{0}$ is always feasible.
Note that $t_i$ represents the distance to point $i$'s nearest neighbor since constraints \eqref{eq:big_lp_dual_constr_t} ensure that $t_i \leq \min_{j \in \mc{C}_i} d_{ij}$,
while the objective function encourages $t_i = \min_{j \in \mc{C}_i} d_{ij}$ for all $i$.
Meanwhile, constraints \eqref{eq:big_lp_separation_constr_t} ensure that the separation $[\min_{k \in \mc{N}\backslash\mc{C}_i} d_{ik}] - [\min_{j \in \mc{C}_i} d_{ij}]$ is at least $\epsilon$.
If model \eqref{model:big_lp_1} returns a feasible solution with $t_i^* = \min_{j \in \mc{C}_i} d_{ij}^* > 0$ for all $i \in \mc{N}$, then the distance metric satisfies requirements 1-3.
Otherwise there exists a subset $\mc{O}$ of outlier points such that $0 \leq t_i^* < \min_{j \in \mc{C}_i} d_{ij}^*$ for all $i \in \mc{O}$, which means that no such distance metric guaranteeing a minimum separation exists for the choice of $\epsilon$.

The following algorithm can be used to explore potential distance metrics for different levels of separation.

\begin{algorithm}
%{\small
\caption{Iterative LP Heuristic}
\label{algo:iterative_LP_heuristic}
\begin{algorithmic}[1]
\STATE Solve model \eqref{model:epsilon_lp_1}. Let $\epsilon^{\max}$ be the optimal objective function value.
\STATE Let $\mc{E} = \{ i / \epsilon^{\max} : i=1,\ldots,I \}$ (e.g., $I=10$)
\FOR{ $\epsilon \in \mc{E}$ }
	\STATE Solve model \eqref{model:big_lp_1}.
	\STATE Store the outliers and distance metric
\ENDFOR
\end{algorithmic}
%}
\end{algorithm}

\newpage
\section{Mixed-Integer Linear Optimization Approaches}

\subsection{Max-Min Approach: Maximize the minimum separation with outliers}

Model \eqref{model:1} can be expressed as a huge mixed-integer linear programming problem:
\begin{subequations} \label{model:milp_with_outlier_detection}
\begin{alignat}{4}
\max_{\lambda,\v{B},\v{d},\v{w},\v{y}}~~& f(\vlambda) = \sum_{i \in \mc{N}} f_i(\lambda_i) = \sum_{i \in \mc{N}} (\lambda_i - \rho z_i) & & \\
\st~~& \sum_{j \in \mc{C}_i} w_{ij} + \lambda_i \leq d_{ik} + M_i z_i  && \qquad \forall i \in \mc{N}, k \in \mc{N}\backslash\mc{C}_i \\
    & \lambda_i \leq 1 - z_i && \qquad \forall i \in \mc{N} \label{eq:outlier_separation_logic} \\
    & w_{ij} \leq d_{ij} && \qquad \forall i \in \mc{N}, i \in \mc{C}_i \label{eq:mccormick_1} \\
    & w_{ij} \leq y_{ij} && \qquad \forall i \in \mc{N}, i \in \mc{C}_i \label{eq:mccormick_2} \\
    & w_{ij} \geq d_{ij} + y_{ij} - 1 && \qquad \forall i \in \mc{N}, i \in \mc{C}_i \label{eq:mccormick_3} \\
    & \sum_{j \in \mc{C}_i} y_{ij} = 1 && \qquad \forall i \in \mc{N} \\
    & d_{ij} = \sum_{k=1}^p \delta_{ijk}^2 b_{kk} + 2 \sum_{k=1}^{p-1} \sum_{\ell=k+1}^p  \delta_{ijk} \delta_{ij\ell} b_{k\ell} && \qquad \forall i,j \in \mc{N} \label{eq:big_milp_distance_def_constr} \\
    & b_{k\ell} \in \Re && \qquad \forall k=1,\dots,p, \ell=k,\dots,p \\
    & d_{ij} \geq 0 && \qquad \forall i,j \in \mc{N} \\
    & w_{ij} \geq 0 && \qquad \forall i \in \mc{N}, i \in \mc{C}_i \\
    & y_{ij} \in \{0,1\} && \qquad \forall i \in \mc{N}, i \in \mc{C}_i \\
    & z_{i} \in \{0,1\} && \qquad \forall i \in \mc{N} \\
    & \lambda_i \geq 0 & & \qquad \forall i \in \mc{N}
\end{alignat}
\end{subequations}

Model \eqref{model:milp_with_outlier_detection} assumes a linear reward for positive separation and a fixed (constant) penalty $\rho$ for each outlier.  
Outliers are handled through the binary $z_{i}$ decision variables.
If $z_{i} = 1$, then constraint \eqref{eq:outlier_separation_logic} ensures that the separation $\lambda_i$ is 0 and hence ignored; otherwise (if $z_{i} = 0$), this constraint is effectively ignored. 
Constraints \eqref{eq:mccormick_1}-\eqref{eq:mccormick_3} are the so-called McCormick envelopes needed to linearize the bilinear equation $w_{ij} = d_{ij} y_{ij}$.

TODO: Modify to incorporate assumption that $0 \leq d_{ij} \leq 1$ for all $i$ and $j$.

\newpage
\section{Generalized disjunctive programming (GDP) formulations}

This section presents several generalized disjunctive programming formulations.

\textbf{Robust GDP formulation with no outliers}
\begin{subequations} \label{model:gdp_robust_no_outliers}
\begin{alignat}{4}
\max_{\lambda,\v{a},\v{B},\v{d}}~~& \lambda &&  \\
\st~~& \mathop{\Xor}_{j \in \mc{C}_i} [ d_{ij} + \lambda \leq d_{ik} \quad \forall k \notin \mc{C}_i ] && \qquad \forall i \in \mc{N} \label{eq:gdp_separation_constr} \\
    & (\v{a},\v{B},\v{d}) \in \mc{F} && 
\end{alignat}
\end{subequations}
Constraints \eqref{eq:gdp_separation_constr} read: ``the distance between point $i$ and its in-class neighbor $j_1 \in \mc{C}_i$ plus the separation $\lambda$ must not exceed the distance between point $i$ and all of its non-neighbors'' or ``the distance between point $i$ and its in-class neighbor $j_2 \in \mc{C}_i$ plus $\lambda$ must not exceed ... '' or ...

\textbf{Robust GDP formulation with outliers}

\begin{subequations} \label{model:gdp_robust_with_outliers}
\begin{alignat}{4}
\max_{\lambda,\v{a},\v{B},\v{d}}~~& \lambda - \rho \sum_{i \in \mc{N}} z_i &&  \\
\st~~& \Bigg[ \mathop{\Xor}_{j \in \mc{C}_i} [ d_{ij} + \lambda \leq d_{ik} \quad \forall k \notin \mc{C}_i ] \Bigg]~\mathop{\Xor}~[z_i = 1] && \qquad \forall i \in \mc{N} \label{eq:gdp_robust_with_outliers_separation_constr} \\
    & (\v{a},\v{B},\v{d}) \in \mc{F}, \v{z} \in \Re_+^N && 
\end{alignat}
\end{subequations}

\textbf{GDP formulation with outliers}

\begin{subequations} \label{model:gdp_nonrobust_with_outliers}
\begin{alignat}{4}
\max_{\vlambda,\v{a},\v{B},\v{d}}~~& \sum_{i \in \mc{N}} \lambda_i &&  \\
\st~~& \Bigg[ \mathop{\Xor}_{j \in \mc{C}_i} [ d_{ij} + \lambda_i \leq d_{ik} \quad \forall k \notin \mc{C}_i; \lambda_i \geq 0 ] \Bigg]~\mathop{\Xor}~[\lambda_i = -\rho] && \qquad \forall i \in \mc{N} \label{eq:gdp_nonrobust_with_outliers_separation_constr} \\
    & (\v{a},\v{B},\v{d}) \in \mc{F}, \lambda_i \geq -\rho && 
\end{alignat}
\end{subequations}

\newpage
\section{Nomenclature}
\begin{center}
\begin{tabular}{cl}
Decision variable & Definition \\
\hline
$d_{ij}$ & distance between points $i$ and $j$ \\
$w_{ij}$ & $d_{ij}y_{ij}$  \\
$y_{ij}$ & 1 if point $j$ is the closest neighbor to $i$; 0 otherwise  \\
$z_{i}$  & 1 if point $i$ is an outlier; 0 otherwise  \\
$\lambda_i$ & separation (difference) between closest non-neighbor and closest neighbor \\$\lambda_i$ & separation (difference) between closest non-neighbor and closest neighbor \\
$\v{B}$ & $\frac{p(p-1)}{2}$ \\ 
\end{tabular}
\end{center}

\begin{proposition}
Note that we may always assume that the data (input vectors) $v_{i}$ have been normalized to reside in the unit hypercube as follows: $v_{ij} \leftarrow (v_{ij} - m)/(M-m)$ where $m=\min\{v_{ij} : i \in \mc{N}, j=1,\dots,D \}$ and $M=\max\{v_{ij} : i \in \mc{N}, j=1,\dots,D \}$.
\end{proposition}

\end{document}